Bayesian Nonparametric System Reliability\\ using Sets of Priors

### 2016-01-11 ijar-npb-ms.tex
Section 3 language tweaks -- feel free to roll back any/all of them!

In mathematical theory of reliability, the main focus is on the functioning of a system given the functioning, or not, 
of its components and the structure of the system. The mathematical concept which is central to this theory is the 
\emph{structure function} \citep{BP75}. For a system with $m$ components, let state vector 
$\underline{x} = (x_1,x_2,\ldots,x_m) \in \{0,1\}^m$, with $x_i=1$ if the $i$th component functions 
and $x_i=0$ if not. The labelling of the components is arbitrary but must be fixed to define $\underline{x}$. 
The structure function $\phi : \{0,1\}^m \rightarrow \{0,1\}$, defined for all possible $\underline{x}$, takes 
the value 1 if the system functions and 0 if the system does not function for state vector $\underline{x}$. 
Most practical systems are coherent, which means that $\phi(\underline{x})$ 
is non-decreasing in any of the components of $\underline{x}$, so system functioning cannot be improved by worse performance 
of one or more of its components. The assumption of coherent systems is also convenient from the perspective of uncertainty
quantification for system reliability. It is further logical to assume that $\phi(\underline{0})=0$ and $\phi(\underline{1})=1$, 
so the system fails if all its components fail and it functions if all its components function.

@@

For each time point $t$, the functioning of a single component of type $k$ at time $t$ (functioning: 1, failed: 0)
is Bernoulli distributed with a suitable functioning probability (as, $1-$failure probability) $\ptk$, so
\begin{align*}
P\big(\indic(T^k_i > t) = 1\big) &= \ptk\,, \\
P\big(\indic(T^k_i > t) = 0\big) &= 1 - \ptk\,,
\end{align*}
in short, $\indic(T^k_i > t) \sim \ber(\ptk)$, $i = 1, \ldots, m_k$, $t \in {\cal T}$.

The set of functioning probabilities $\{ \ptk, t \in {\cal T}\}$
defines a discrete failure time distribution for components of type $k$ through
\begin{align*}
R^k(t_j) &= P(T^k > t_j) = p^k_{t_j},\ t_j = t_1, \ldots, \tmax\,.
\end{align*}
We can express this failure time distribution also through the probability mass function (pmf) and the discrete hazard function,
\begin{align*}
f^k(t_j) &= P\big(T^k \in (t_j,t_{j+1}]\big) = p^k_{t_j} - p^k_{t_{j+1}}\,,\\ 
h^k(t_j) &= P\big(T^k \in (t_j,t_{j+1}]\mid T^k > t_j\big) = \frac{f^k(t_j)}{R^k(t_j)}\,. % or R^k(t_{j-1}) ???
\end{align*}
We assume the time grid $\cal T$ is dense enough for the application at hand,
otherwise it would be possible to assume, e.g.,
$R^k(t) = p^k_{t_j}, t \in [t_j, t_{j+1})$,
or taking $p^k_{t_j}$ and $p^k_{t_{j+1}}$ as upper and lower bounds for $R^k(t)$, $t \in [t_j, t_{j+1})$.

@@

Due to the independence assumption for components of the same type, 
the number of functioning components out of the $m_k$ components of type $k$
is binomially distributed, $C^k_t = \sum_{i=1}^{m_k} \indic(T^k_i > t) \sim \bin(\ptk, m_k)$.
%(maybe not relevant for this manuscript, but would it make sense to take Kaplan-Meier estimates for the $p^k_t$'s,
%or the NPI bounds for Kaplan-Meier?)

The $\ptk$'s can be taken such that they reflect, e.g., a bathtub curve for the corresponding hazard rate $h^k(t_j)$.
Naturally, $p^k_{t_j} \ge p^k_{t_{j+1}}$ should hold (assuming no repair).

However, directly choosing the $\ptk$'s is hard,
and fixing values for all $p^k_t$ would furthermore neglect any inherent uncertainty in the particular choice.  
To account for this uncertainty, one can express knowledge on $\ptk$ through a prior distribution.
A convenient and natural choice is $\ptk \sim \be(\alpha^k_t, \beta^k_t)$, a conjugate prior,
leading to the posterior being again Beta.

Having observed the lifetime data $\vec{t}^k = (t^k_1, \ldots, t^k_{n_k})$,
at any fixed time $t \in {\cal T}$ this results in the observation from the Binomial model described above,
$s^k_t = \sum_{i=1}^{n_k} \indic(t^k_i > t)$.
The posterior is then $\ptk \mid s^k_t \sim \be(\alpha^k_t + s^k_t, \beta^k_t + n_k - s^k_t)$.
The combination of a Binomial observation model with a Beta prior is often called Beta-Binomial model.

The posterior predictive distribution for the number of components surviving at time $t$
of the $m_k$ components in the system, based on the lifetime data and the prior information,
is then a so-called Beta-Binomial distribution,
$C^k_t \mid s^k_t \sim \bebin(m_k, \alpha^k_t + s^k_t, \beta^k_t + n_k - s^k_t)$.
That is, we have
\begin{align*}
P(C^k_t = l_k \mid s^k_t) &= {m_k \choose l_k} \frac{B(l_k + \alpha^k_t + s^k_t, m_k - l_k + \beta^k_t + n_k - s^k_t)}
                                                    {B(\alpha^k_t + s^k_t, \beta^k_t + n_k - s^k_t)} \,,
\end{align*}
where $B(\cdot, \cdot)$ is the Beta function.
This posterior predictive distribution is basically a Binomial distribution
where the functioning probability $\ptk$ takes values in $[0,1]$
with the probability given by the posterior on $\ptk$.

To wrap up, to calculate the system reliability according to \eqref{eq:rsyswithsurvsign},
for each component type $k$,
we need lifetime data $\vec{t}^k$,
and have to choose $2 \times |{\cal T}|$ parameters
to specify the prior distribution for the discrete survival function of $T^k_i$.
In total, values for $2 \times |{\cal T}| \times K$ parameters must be chosen.

>>>>

In the mathematical theory of reliability, the main focus is on the functioning of a system given the functioning, or not, 
of its components and the structure of the system. The mathematical concept which is central to this theory is the 
\emph{structure function} \citep{BP75}. For a system with $m$ components, let state vector 
$\underline{x} = (x_1,x_2,\ldots,x_m) \in \{0,1\}^m$, with $x_i=1$ if the $i$th component functions 
and $x_i=0$ if not. The labelling of the components is arbitrary but must be fixed to define $\underline{x}$. 
The structure function $\phi : \{0,1\}^m \rightarrow \{0,1\}$, defined for all possible $\underline{x}$, takes 
the value 1 if the system functions and 0 if the system does not function for state vector $\underline{x}$. 
Most practical systems are coherent, which means that $\phi(\underline{x})$ 
is non-decreasing in any of the components of $\underline{x}$, so system functioning cannot be improved by worse performance 
of one or more of its components. The assumption of coherent systems is also convenient from the perspective of uncertainty
quantification for system reliability. It is further logical to assume that $\phi(\underline{0})=0$ and $\phi(\underline{1})=1$, 
so the system fails if all its components fail and it functions if all its components function.

@@

At each time point $t$, the operational state of a single component of type $k$
is Bernoulli distributed (functioning: 1, failed: 0) with parameter $\ptk$, so that
\begin{align*}
P\big(\indic(T^k_i > t) = 1\big) &= \ptk\,, \\
P\big(\indic(T^k_i > t) = 0\big) &= 1 - \ptk\,,
\end{align*}
That is, $\indic(T^k_i > t) \sim \ber(\ptk)$, $i = 1, \ldots, m_k$, $t \in {\cal T}$.

The set of probabilities $\{ \ptk, t \in {\cal T}\}$
defines a discrete failure time distribution for components of type $k$ through
\begin{align*}
R^k(t_j) &= P(T^k > t_j) = p^k_{t_j},\ t_j = t_1, \ldots, \tmax\,.
\end{align*}
We can also express this failure time distribution through the probability mass function (pmf) and discrete hazard function,
\begin{align*}
f^k(t_j) &= P\big(T^k \in (t_j,t_{j+1}]\big) = p^k_{t_j} - p^k_{t_{j+1}}\,,\\ 
h^k(t_j) &= P\big(T^k \in (t_j,t_{j+1}]\mid T^k > t_j\big) = \frac{f^k(t_j)}{R^k(t_j)}\,. % or R^k(t_{j-1}) ???
\end{align*}
The time grid $\cal T$ can be chosen to be appropriately dense for the application at hand,
with the natural extension between grid points by taking $R^k(\cdot)$ to be the right continuous step function induced by the grid values,
$R^k(t) = p^k_{t_j}, t \in [t_j, t_{j+1})$,
or by taking $p^k_{t_j}$ and $p^k_{t_{j+1}}$ as upper and lower bounds for $R^k(t)$, $t \in [t_j, t_{j+1})$.

@@

The independence assumption for components of the same type immediately implies that 
the number of functioning components of type $k$ in the system
is binomially distributed, $C^k_t = \sum_{i=1}^{m_k} \indic(T^k_i > t) \sim \bin(\ptk, m_k)$.
%(maybe not relevant for this manuscript, but would it make sense to take Kaplan-Meier estimates for the $p^k_t$'s,
%or the NPI bounds for Kaplan-Meier?)

The $\ptk$'s can, in theory, be directly chosen to arbitrarily closely approximate any valid lifetime pdf on
$[0,\infty)$, for example matching a bathtub curve for the corresponding hazard rate $h^k(t_j)$.
Naturally, $p^k_{t_j} \ge p^k_{t_{j+1}}$ should hold (assuming no repair).
However, such direct specification is non-trivial, neglects any inherent uncertainty in the particular choice, and cannot be 
easily combined with test data.
To account for the uncertainty, one can express knowledge about $\ptk$ through a prior distribution.
A convenient and natural choice is $\ptk \sim \be(\alpha^k_t, \beta^k_t)$, particularly because in a Bayesian inferential
setting this is the conjugate prior which leads to a Beta posterior.

Let the lifetime test data collected on component $k$ be $\vec{t}^k = (t^k_1, \ldots, t^k_{n_k})$.
At each fixed time $t \in {\cal T}$, this corresponds to an observation from the Binomial model described above,
$s^k_t = \sum_{i=1}^{n_k} \indic(t^k_i > t)$.
The posterior is then $\ptk \mid s^k_t \sim \be(\alpha^k_t + s^k_t, \beta^k_t + n_k - s^k_t)$.
The combination of a Binomial observation model with a Beta prior is often called Beta-Binomial model.

The posterior predictive distribution for the number of components surviving at time $t$
in a new system, based on the lifetime data and the prior information,
is then a so-called Beta-Binomial distribution,
$C^k_t \mid s^k_t \sim \bebin(m_k, \alpha^k_t + s^k_t, \beta^k_t + n_k - s^k_t)$.
That is, we have
\begin{align*}
P(C^k_t = l_k \mid s^k_t) &= {m_k \choose l_k} \frac{B(l_k + \alpha^k_t + s^k_t, m_k - l_k + \beta^k_t + n_k - s^k_t)}
                                                    {B(\alpha^k_t + s^k_t, \beta^k_t + n_k - s^k_t)} \,,
\end{align*}
where $B(\cdot, \cdot)$ is the Beta function.
This posterior predictive distribution is essentially a Binomial distribution
where the functioning probability $\ptk$ takes values in $[0,1]$
with the probability given by the posterior on $\ptk$.

Consequently, in order to calculate the system reliability according to \eqref{eq:rsyswithsurvsign},
for each component type $k$
we need lifetime data $\vec{t}^k$,
and have to choose $2 \times |{\cal T}|$ parameters
to specify the prior distribution for the discrete survival function of $T^k_i$.
In total, values for $2 \times |{\cal T}| \times K$ parameters must be chosen.
